Large-scale hybrid ad hoc network for mobile platforms: Challenges and Experiences

### 2019-02-11 paper.tex
Finalized algorithms

The following describes how the c2p2p architecture addresses the
challenges identified above along with known limitations. Further, to
realize this architecture on mobile devices, we evaluated a variety of
protocols on both Android and iOS. As described later, WiFi DNS on
Android, Bluetooth Nearby on Android, and Blutooth low energy (BTLE)
on both Android and iOS are the only protocols that were found
effective. Actual techniques for discovery, advertisement, and
connection vary across these protocols and are described next.
%
\subsection{Operating system restrictions}
\label{sec:os}
MNA works around the background activity and user permission
restrictions via innovative techniques, without resorting to hacks
that may violate user security or App-store guidelines.
Tables~\ref{tab:android} and~\ref{tab:ios} summarize the various
protocols we have experimented with using WiFi and Bluetooth
interfaces, along with key findings as described here.  The protocols
actually deployed are in bold.
%
%
\begin{table}[H]
\caption{Android protocols and key findings}
\label{tab:android}
\centering
\begin{tabular}{|c|l|r|r|r|c|l|}
\hline
& \multicolumn{1}{c|}{Protocol} & \multicolumn{1}{c|}{Throughput} & \multicolumn{1}{c|}{Battery} & \multicolumn{1}{c|}{Range} & \multicolumn{1}{c|}{iOS} & \multicolumn{1}{c|}{Issues}\\
&                               & \multicolumn{1}{c|}{Avg. kbps}  & \multicolumn{1}{c|}{per hour} & \multicolumn{1}{c|}{LoS ft} &                  &                            \\
\hline
\parbox[t]{2mm}{\multirow{3}{*}{\rotatebox[origin=c]{90}{WiFi}}} & \textbf{DNS} & 0.2 & 2\% & 600 & No & \\
& WiDi & 2000 & 3\% & 600 & No & Needs DNS\\
& Hotspot & 2000 & 5\% & 600 & No & Permission\\
\hline
\parbox[t]{2mm}{\multirow{3}{*}{\rotatebox[origin=c]{90}{BT}}} & Classic & 50 & 2\% & 600 & No & Unstable\\
& \textbf{Nearby} & 50 & 2\% & 600 & No & \\
& \textbf{BTLE} & 50 & 2\% & 600 & Yes & \\
\hline
\end{tabular}
\end{table}
%
%
\begin{table}[H]
\caption{iOS protocols and key findings}
\label{tab:ios}
\centering
\begin{tabular}{|c|l|r|r|r|c|l|}
\hline
& \multicolumn{1}{c|}{Protocol} & \multicolumn{1}{c|}{Throughput} & \multicolumn{1}{c|}{Battery} & \multicolumn{1}{c|}{Range} & \multicolumn{1}{c|}{Android} & \multicolumn{1}{c|}{Issues}\\
&                               & \multicolumn{1}{c|}{Avg. kbps}  & \multicolumn{1}{c|}{per hour} & \multicolumn{1}{c|}{LoS ft} &                  &                            \\
\hline
\parbox[t]{2mm}{\multirow{2}{*}{\rotatebox[origin=c]{90}{WiFi}}} & Bonjour & 0.2 & 5\% & 200 & No & Battery\\
& MPC & 2000 & 5\% & 200 & No & Battery\\
\hline
& \textbf{BTLE} & 50 & 1\% & 800 & Yes & \\
\hline
\end{tabular}
\end{table}
%

@@

%
\subsection{Trust in information}
\label{sec:trust}
%
In our approach, since the origin of weather information is the
Weather Channel service, digital signatures can be attached to all
weather alerts broadcast from the service. The mobile application is
distributed with the corresponding public key so that digital
signatures from the service can be verified. If a message fails such a
verification, it is discarded and not forwarded any further.
%
\section{Algorithms}
\label{sec:algo}
%
\subsection{WiFi DNS}
\label{sec:wifi}
%
DNS service discovery is a widely supported
protocol \cite{cheshire-dns-sd-2013}. On Android, WiFi DNS improvises
on this standard such that the exchange of information happens without
making network connections at all. This is achieved by splitting the
message payloads into small enough chunks and stuffing them into
service advertisements as txt records and broadcast over multiple
advertisements in a quick succession. Since there are no connections
being made, the messages can be exchanged without prompting the user
to grant permissions for each connection. Since no peer connections
are needed and advertisements and discovery are broadcast, WiFi DNS is
a broadcast protocol. On iOS, Bonjour protocol is built on DNS service
discovery but is power-hungry and does not interoperate with WiFi DNS
on Android. Algorithm~\ref{algo:dns} outlines the peer process for
WiFi DNS.

\begin{algorithm}[htb!]
\label{algo:dns}
\DontPrintSemicolon % Some LaTeX compilers require you to use \dontprintsemicolon instead
\SetAlgoLined
\KwIn{Peer id $p_{id}$}
\KwIn{App id $a_{id}$, max advert/discovery intervals $max_{ad}$, $max_{disc}$}
\KwIn{min advert/discovery intervals $min_{ad}$, $min_{disc}$}
 $S_{name} \gets$ \textsf{'MNA'}\;
 $\mathbb{M} \gets \varnothing$\;
 $T_{ad} \gets min_{ad}$\;
 $T_{disc} \gets min_{disc}$\;
 $New \gets 0$
 \ForEach{timer expiry $T_{ad}$}{
   $\mathbb{M} \gets \mathbb{M} -$ expired messages\;
   $\mathbb{M}_{sort} \gets$ $\mathbb{M}$ sorted by least advertised first\;
   $B \gets$ $\mathbb{M}_{sort}$ chunked by MTU size\;
   dnsAdvertise ($S_{name}$, $B$)\;
   \If{$New = 0$}{
     $T_{ad} \gets min(T_{ad} \times 2, max_{ad})$\;
   }
 }
 \ForEach{timer expiry $T_{disc}$}{
   dnsDiscover ($S_{name}$)\;
   \If{$New = 0$}{
     $T_{disc} \gets min(T_{disc} \times 2, max_{disc})$\;
   }
   $New \gets 0$\;
 }
 \ForEach{discovered service callback $S_{disc}$}{
   $B \gets S_{disc}.B$\;
   $M_{part} \gets msg(B)$\;
   $M \gets M + M_{part}$\; 
   \If{$M_{part}$ was the last chunk of $M$}{
     \If{$M$ is not a duplicate}{
       \If{$M$ is unexpired $\wedge$ $M$ has valid signature}{
         $\mathbb{M} \gets \mathbb{M} \cup M$\;
         $T_{ad} \gets max(\frac{T_{ad}}{2}, min_{ad})$\;
         $T_{disc} \gets max(\frac{T_{disc}}{2}, min_{disc})$\;
         $New \gets 1$\;
         \If{$M.app = a_{id}$}{
           Notify app with $M$\;
         }
       }
     }
   }
 }
 \tcc{For testing only}
 \ForEach{app request to send $M$}{
   \If{$M$ has not expired}{
     $\mathbb{M} \gets \mathbb{M} \cup M$\;
   }
 }
 \caption{WiFi DNS peer algorithm}
\end{algorithm}
%
\subsection{BTLE}
\label{sec:btle}
%
The only protocol that stays reliably active in
background without violating iOS developer guidelines is BTLE. Unlike
Android, iOS applications do not need to request discovery and
advertisements contunually. Once a BTLE Peripheral (role that has
information) is advertised and a Central (role that receives
information) requests discovery, the app is allowed to go to
sleep. When a matching central or a peripheral device is found, iOS
wakes up the application app to handle such an event, even when the
application is in the background. MNA builds on this and leverages
acynchronous dispatch queues to turn each peer into a Central as well
as a Peripheral simultaneously. This architecture allows bidirectional
data transfers without concurrency issues.

>>>>

The following describes how the c2p2p architecture addresses the
challenges identified above along with known limitations. Further, to
realize this architecture on mobile devices, we evaluated a variety of
protocols on both Android and iOS. As described later, WiFi DNS on
Android, Bluetooth Nearby on Android, and Blutooth low energy (BTLE)
on both Android and iOS are the only protocols that were found
effective. Actual techniques for discovery, advertisement, and
connection vary across these protocols and are described next.
%
\subsection{Operating system restrictions}
\label{sec:os}
MNA works around the background activity and user permission
restrictions via innovative techniques, without resorting to hacks
that may violate user security or App-store guidelines.
Tables~\ref{tab:android} and~\ref{tab:ios} summarize the various
protocols we have experimented with using WiFi and Bluetooth
interfaces, along with key findings as described here.  The protocols
actually deployed are in bold.
%
%
\begin{table}[H]
\caption{Android protocols and key findings}
\label{tab:android}
\centering
\begin{tabular}{|c|l|r|r|r|c|l|}
\hline
& \multicolumn{1}{c|}{Protocol} & \multicolumn{1}{c|}{Throughput} & \multicolumn{1}{c|}{Battery} & \multicolumn{1}{c|}{Range} & \multicolumn{1}{c|}{iOS} & \multicolumn{1}{c|}{Issues}\\
&                               & \multicolumn{1}{c|}{Avg. kbps}  & \multicolumn{1}{c|}{per hour} & \multicolumn{1}{c|}{LoS ft} &                  &                            \\
\hline
\parbox[t]{2mm}{\multirow{3}{*}{\rotatebox[origin=c]{90}{WiFi}}} & \textbf{DNS} & 0.2 & 2\% & 600 & No & \\
& WiDi & 2000 & 3\% & 600 & No & Needs DNS\\
& Hotspot & 2000 & 5\% & 600 & No & Permission\\
\hline
\parbox[t]{2mm}{\multirow{3}{*}{\rotatebox[origin=c]{90}{BT}}} & \textbf{Classic} & 50 & 2\% & 600 & No & \\
& \textbf{Nearby} & 50 & 2\% & 600 & No & \\
& \textbf{BTLE} & 50 & 2\% & 600 & Yes & \\
\hline
\end{tabular}
\end{table}
%
%
\begin{table}[H]
\caption{iOS protocols and key findings}
\label{tab:ios}
\centering
\begin{tabular}{|c|l|r|r|r|c|l|}
\hline
& \multicolumn{1}{c|}{Protocol} & \multicolumn{1}{c|}{Throughput} & \multicolumn{1}{c|}{Battery} & \multicolumn{1}{c|}{Range} & \multicolumn{1}{c|}{Android} & \multicolumn{1}{c|}{Issues}\\
&                               & \multicolumn{1}{c|}{Avg. kbps}  & \multicolumn{1}{c|}{per hour} & \multicolumn{1}{c|}{LoS ft} &                  &                            \\
\hline
\parbox[t]{2mm}{\multirow{2}{*}{\rotatebox[origin=c]{90}{WiFi}}} & Bonjour & 0.2 & 5\% & 200 & No & Battery\\
& MPC & 2000 & 5\% & 200 & No & Battery\\
\hline
& \textbf{BTLE} & 50 & 1\% & 800 & Yes & \\
\hline
\end{tabular}
\end{table}
%

@@

%
\subsection{Trust in information}
\label{sec:trust}
%
In our approach, since the origin of weather information is the
Weather Channel service, digital signatures can be attached to all
weather alerts broadcast from the service. The mobile application is
distributed with the corresponding public key so that digital
signatures from the service can be verified. If a message fails such a
verification, it is discarded and not forwarded any further.
%
\section{Algorithms}
\label{sec:algo}
%
\subsection{WiFi DNS}
\label{sec:wifi}
%
DNS service discovery is a widely supported protocol
\cite{cheshire-dns-sd-2013}. On Android, WiFi DNS improvises on this
standard such that the exchange of information happens without making
network connections at all. This is achieved by splitting the message
payloads into small enough chunks and stuffing them into service
advertisements as txt records and broadcast over multiple
advertisements in a quick succession. Since there are no connections
being made, the messages can be exchanged without prompting the user
to grant permissions for each connection. Since no peer connections
are needed and advertisements and discovery are broadcast, WiFi DNS is
a broadcast protocol. On iOS, Bonjour protocol is built on DNS service
discovery but is power-hungry and does not interoperate with WiFi DNS
on Android. Algorithm~\ref{algo:dns} outlines the peer process for
WiFi DNS. Power conservation via time intervals is achieved by
manupulating $T_{ad}$ and $T_{disc}$ based on traffic.

Messages from multiple applications are forwarded on the same network,
but a peer can validate signatures only for the messages targeting an
app it hosts. Peers simply forward messages not targeting one of their
apps. Key benefit of this design is that the density of peers is
available to all apps, making c2p2p scenarios viable even when an
individual app does not have a sufficient density of users.  However,
this also allows a malicioous jammer to generate traffic for a fake
app and force the peers to forwad it (as they do not match any of the
apps) until the messages expire.  Although the fake information will
never be delivered to the apps, this issue can affect MNA throughput
for legitimate apps. A possible solution would be to maintain a
central registry of $a_{id}$ and distribute it to all peers across
apps.

\begin{algorithm}[htb!]
\label{algo:dns}
\DontPrintSemicolon % Some LaTeX compilers require you to use \dontprintsemicolon instead
\SetAlgoLined
\KwIn{Peer id $p_{id}$}
\KwIn{App id $a_{id}$, max advert/discovery intervals $max_{ad}$, $max_{disc}$}
\KwIn{min advert/discovery intervals $min_{ad}$, $min_{disc}$}
 $S_{name} \gets$ \textsf{'MNA'}\;
 $\mathbb{M} \gets \varnothing$\;
 $T_{ad} \gets min_{ad}$\;
 $T_{disc} \gets min_{disc}$\;
 $New \gets 0$
 \ForEach{timer expiry $T_{ad}$}{
   $\mathbb{M} \gets \mathbb{M} -$ expired messages\;
   $\mathbb{M}_{sort} \gets$ $\mathbb{M}$ sorted by least advertised first\;
   $B \gets$ $\mathbb{M}_{sort}$ chunked by MTU size\;
   dnsAdvertise ($S_{name}$, $p_{id}$, $B$)\;
   \If{$New = 0$}{
     $T_{ad} \gets min(T_{ad} \times 2, max_{ad})$\;
   }
 }
 \ForEach{timer expiry $T_{disc}$}{
   dnsDiscover ($S_{name}$, $p_{id}$)\;
   \If{$New = 0$}{
     $T_{disc} \gets min(T_{disc} \times 2, max_{disc})$\;
   }
   $New \gets 0$\;
 }
 \ForEach{discovered service callback $S_{disc}$}{
   $B \gets S_{disc}.B$\;
   $M_{part} \gets msg(B)$\;
   $M \gets M + M_{part}$\; 
   \If{$M_{part}$ is not a duplicate}{
        $T_{ad} \gets max(\frac{T_{ad}}{2}, min_{ad})$\;
        $T_{disc} \gets max(\frac{T_{disc}}{2}, min_{disc})$\;
        $New \gets 1$\;
   }
   \If{$M_{part}$ was the last chunk of $M$}{
     \If{$M$ is unexpired}{
         \If{$M.app = a_{id}$ $\wedge$ $M$ signature is valid}{
             Notify app with $M$\;
         }
         $\mathbb{M} \gets \mathbb{M} \cup M$\;
     }
   }
 }
 \tcc{For testing only}
 \ForEach{app request to send $M$}{
   \If{$M$ has not expired}{
     $\mathbb{M} \gets \mathbb{M} \cup M$\;
   }
 }
 \caption{WiFi DNS peer algorithm}
\end{algorithm}
%
\subsection{BTLE}
\label{sec:btle}
%
The only protocol that stays reliably active in
background without violating iOS developer guidelines is BTLE. Unlike
Android, iOS applications do not need to request discovery and
advertisements contunually. Once a BTLE Peripheral (role that has
information) is advertised and a Central (role that receives
information) requests discovery, the app is allowed to go to
sleep. When a matching central or a peripheral device is found, iOS
wakes up the application app to handle such an event, even when the
application is in the background. MNA builds on this and leverages
acynchronous dispatch queues to turn each peer into a Central as well
as a Peripheral simultaneously. This architecture allows bidirectional
data transfers without concurrency issues.

\begin{algorithm}[htb!]
\label{algo:btle}
\DontPrintSemicolon % Some LaTeX compilers require you to use \dontprintsemicolon instead
\SetAlgoLined
\KwIn{Peer id $p_{id}$, App id $a_{id}$, Max peripherals $max_p$}
 $S_{name} \gets$ \textsf{'MNA'}\;
 $\mathbb{M} \gets \varnothing$\;
 $C \gets$ new Central() in new thread\;
 $P \gets$ new Peripheral() in new thread\;
 $T \gets$ new data sender thread\;
 $C$.bleDiscovery($p_{id}$, $S_{name}$)\;
 $P$.bleAdvert($p_{id}$, $S_{name}$)\;
 $P_{con} \gets 0$\;

 \ForEach{Peripheral $peer$ found event}{
   \If{$P_{con} < max_p$}{
     $C$.connect($peer$)\;
     $P_{con} \gets P_{con} + 1$\;
   }
 }

 \ForEach{Peripheral $peer$ connected event}{
   \ForEach{$M$ $\in$ $\mathbb{M}$}{
     \If{$peer \notin M$.fwdList}{
       $T$.send($M$, $peer$)\;
     }
   }
 }

 \ForEach{Peripheral $peer$ disconnected event}{
    $P_{con} \gets P_{con} - 1$\;
 }

 \ForEach{Central $peer$ connected event}{
   \ForEach{$M$ $\in$ $\mathbb{M}$}{
     \If{$peer \notin M$.fwdList}{
       $T$.send($M$, $peer$)\;
     }
   }
 }

 \ForEach{$M$ received event from $T$.receive()}{
   \If{$M$ is not a duplicate $\wedge$ $M$ is unexpired}{
     \If{$M.app = a_{id}$ $\wedge$ $M$ signature is valid}{
       Notify app with $M$\;
     }
     $M$.fwdList $\gets$ $M$.fwdList $\cup$ $p_{id}$\;
     $\mathbb{M} \gets \mathbb{M} \cup M$\;     
     $T$.send($M$, all connected peers)\;
   }
 }
 \tcc{For testing only}
 \ForEach{app request to send $M$}{
   \If{$M$ has not expired}{
     $M$.fwdList $\gets$ $p_{id}$ \;
     $\mathbb{M} \gets \mathbb{M} \cup M$\;
     $T$.send($M$, all connected peers)\;
   }
 }
 \caption{BTLE peer algorithm}
\end{algorithm}
